---
title: "R Notebook"
output: html_notebook
---
#Z. Noltii Field Biomass (initial)
#JW June 2025



```{r}
library(ggplot2)
library(tidyverse)
library(readr)
```



#Lab_Cores.csv  (from Z. Noltii Google Drive /Raw Data/Field/Field_Ambient_Seagrass sheet -- Lab_Cores) 

#forcats::as_factor ensures order stays the same as inputted into csv (unsure why Biomass_Part utilizes this function)
#mutate(TSC... used to change Total Shoot Count per core (TSC) into TSC per m2
```{r}

ABZ <- read_csv("Data/Lab_Core.csv", na = c("","NA")) 
ABZ$Month <- forcats::as_factor(ABZ$Month)
ABZ$Biomass_Part <- forcats::as_factor(ABZ$Biomass_Part)
ABZ <- ABZ %>% mutate(TSC=(TSC*(100/pi))) %>% mutate(ID=row_number()) %>% rename("SGP"="Biomass_Part", "BDW"="Bio_Dry_Weight_m2") %>%  select(-Processing_Date, -Sample_Date, -Sampler, -Dry_Weight, -Tin_Weight, -Bio_Dry_Weight, -Date.in...out.of.oven, -Notes)
ABZ
```




#Subsets data to only include "Above" (Leaves)
```{r}
AboveZ <- ABZ %>% 
  filter(SGP =="Above")
AboveZ
```


#StatsAZ gives per month mean of Leaf Bio Dry Weight (per m^2) and Standard Error
```{r}
StatsAZ <- AboveZ %>% group_by(Month) %>% summarise(mean= mean(BDW), sd = sd(BDW), n=n(), se=sd/sqrt(n)) %>% mutate(ID=1)
StatsAZ <- as.data.frame(StatsAZ)
StatsAZ
```


#Monthly progression of mean leaf dry weight +- SE
#Y LIMITS MAY NEED TO BE CHANGED AS TIME CONTINUES
```{r}
ggplot(StatsAZ, aes(x=Month, y = mean, group= ID))+

  geom_point()+
  
  geom_errorbar(aes(ymin=mean-se, ymax = mean+se, width=0.2))+
  geom_line()+
  
  labs(y=bquote(~"Leaf"~ (g~m^"-2")))+
  theme_classic()+
  ylim(0,20)

```


#subsets data to only include "Below" (Rhizomes + Roots)
```{r}
BelowZ <- ABZ %>% 
  filter(SGP =="Below")
BelowZ
```

#StatsBZ subsets data for monthly mean below bdw (mutate(ID=1) used to ensure geom_line connects to datapoints)
```{r}
StatsBZ <- BelowZ %>% group_by(Month) %>% summarise(mean= mean(BDW), sd = sd(BDW), n=n(), se=sd/sqrt(n)) %>% mutate(ID=1)
StatsBZ
```

#Monthly progression of mean Roots+Rhizomes +- SE (aes(group=ID) used to ensure geom_line will connect to different points) 
#Y LIMITS MAY NEED TO BE CHANGED AS TIME CONTINUES
```{r}
ggplot(StatsBZ, aes(x=Month, y = mean, group=ID))+
  geom_point()+
  geom_line()+
  geom_errorbar(aes(ymin=mean-se, ymax = mean+se, width=0.2))+
  labs(y=bquote(~"Rhizome & Roots"~(g~m^"-2")))+
  theme_classic()+
  ylim(0,20)
```

#Total Shoot count monthly mean and SE
```{r}
StatsTSC <- ABZ %>% group_by(Month) %>% summarise(mean= mean(TSC, na.rm=TRUE), sd = sd(TSC, na.rm=TRUE), n=n(), se=sd/sqrt(n)) %>%  mutate(ID=1)
StatsTSC
```

#Monthly progression of mean Total Shoot count +- SE
```{r}
ggplot(StatsTSC, aes(x=Month, y=mean, group = ID))+
  geom_point()+
  geom_line()+
  geom_errorbar(aes(ymin=mean-se, ymax = mean+se, width=0.2))+
  labs(y=bquote(~"Shoot Density"~("Shoot"~m^"-2")))+
  theme_classic()
```


#subsets data for only April and acquires mean of Above, Below, and Total BDW and the ratio of leaf to rhizome (Unable to calculate SE as Dry weight Above and Below measured seperately and not tied with Sample ID and Replicate)
```{r}
AprilAB <- subset(ABZ, ABZ$Month == "April") %>% select(-Replicate, -Sample_ID, -TSC) %>%  group_by(SGP)%>% mutate(ID=row_number()) %>%  pivot_wider(names_from = SGP, values_from = BDW) %>% ungroup() %>% select(-"NA") %>% group_by(Month) %>% summarise(Above=mean(Above), Below=mean(Below)) %>% mutate(TotalBDW= Above+Below, ABRat= Above/Below)
AprilAB
```

#adds column names TSE and RSE to row bind AprilTR to ABR2
```{r}
AprilTR <- AprilAB %>% select(-Above, -Below) %>% mutate(TSE=NA, RSE=NA) 
AprilTR
```


#Subset of data for all months OTHER than April, and keeps "Above" and "Below" BDW with respective SampleID and replicate
```{r}
ABR1 <- subset(ABZ, ABZ$Month != "April") %>% select(-TSC, -ID)  %>% pivot_wider( names_from = SGP, values_from = BDW) %>% mutate(SampRep= paste(Sample_ID, Replicate)) %>% mutate(TotalBDW=Above+Below, ABRat= Above/Below) %>% group_by(Month) %>% select(-Sample_ID, -Replicate)
ABR1
```


#Finds mean of Total BDW and SE (TSE) and mean of Leaf to Rhizome ratio (ABRat) and SE (RSE) 
```{r}
ABR2 <-ABR1 %>% summarise(TSE=sd(TotalBDW)/sqrt(n()), RSE=sd(ABRat)/sqrt(n()), TotalBDW=mean(TotalBDW), ABRat= mean(ABRat) )
ABR2
```
#Binds ABR2 and AprilTR together 
```{r}
ABR <- rbind(ABR2, AprilTR) %>% mutate(ID=1)
ABR
```


#Total BDW per month +- SE 
```{r}
ggplot(ABR, aes(x=Month, y=TotalBDW, group=ID))+
  geom_point()+
  geom_line()+
  geom_errorbar(aes(ymax = TotalBDW+TSE, ymin=TotalBDW-TSE, width=0.2))+
  labs(y=bquote(~"Total Weight"~(g~m^"-2")))+
  theme_classic()
```


#Above:Below Ratio per Monthsa

```{r}
ggplot(ABR, aes(x=Month, y=ABRat, group=ID))+
  geom_point()+
  geom_line()+
  geom_errorbar(aes(ymax = ABRat+RSE, ymin=ABRat-RSE, width=0.2))+
  labs(y="A:B")+
  theme_classic()
```